### 2020-12-12 Thesis-Chapter7-PathsAhead.tex
Editorial improvements

The interests of security, safety, privacy, and trust are converging on the issue of encryption and \acl{EA} and
creating a \ac{wicked-prob}. Information security is increasingly important as more data, business, and infrastructure
goes online. The technical community is struggling with security; defending against the average attacker is hard, and
against the best attackers is nearly impossible, even for the most motivated and well-resourced organizations.
\Ac{cryptography} plays a fundamental role in security, yet its simultaneous role in enabling unprecedented privacy
means it has controversial side effects.

One of these side effects is encryption-enabled privacy's impact on public safety. Despite a shortage of quantitative
analyses showing its impact on crime---something that would be impossible to fully quantify, after all---encryption
decidedly withholds evidence of wrongdoing from service providers and law enforcement, enabling truly harmful activity
to go unchecked, and decreasing public safety. This a legitimate concern and the primary motivation for \ac{EA}.

However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
ranging from identity theft, to inappropriate data mining, to \ac{masssurv}. To the extent that encryption mitigates
these, it increases public safety.

Finally, choosing how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
considerable questions of trust. Law enforcement agencies, intelligence agencies, technology makers, individual users,
and the public at large are all players; each have demonstrated untrustworthiness, and different approaches disperse
trust differently.

% TODO: one paragraph here^^ or four?

This complex mix of issues is what makes encryption policy a \ac{wicked-prob}. To merely formulate the problem requires
explanation upon explanation and research upon research. Any formulation is concomitant with a solution, but the outcome
of a solution cannot be accurately forecasted or evaluated in isolation. The underlying conditions are constantly in
flux, affected by external forces as well as every policy action and inaction.

@@

Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption poses a present danger to public safety
in certain respects. The bogeymen of terrorists and criminals using encryption to evade the law are real---in the case
of those peddling and viewing \acl{CSAM}, shockingly so \cite{keller_internet_2019}. However, the argument that
\ac{encryption} is creating circumstances that are to blame for \acl{LE} failure is not currently strong enough to
justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor coordination and training \cite{carter_2018}; for
\ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Additionally, strong \ac{encryption}
plays a key role in protecting security and privacy, public goods that other technological developments often fail to
provide.

Though the pro-\ac{EA} argument is ultimately not strong enough to justify an \ac{EA} mandate, incorporating it to the
technologist's point of view teaches important lessons. One is that wholistic cryptosystem threat models need to include
two non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and
this power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating
steps. Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption;
lawmakers represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too
dangerous. The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.

>>>>

The interests of security, safety, privacy, and trust converge on the issue of encryption and \acl{EA}, creating a
\ac{wicked-prob}. The importance of information security is increasing as more data, business, and infrastructure goes
online. The technical community is struggles with security. Even for the most well-equipped organizations, thwarting the
average attacker is difficult, and the best attacker, impossible. \Ac{cryptography} plays a fundamental role in
security, yet its simultaneous role in enabling unprecedented privacy means it has controversial side effects.

One of these side effects is encryption-enabled privacy's impact on public safety. Despite a shortage of quantitative
analyses showing its impact on crime, encryption certainly does allow evidence of wrongdoing to be hidden from service
providers and law enforcement. This enables harmful activity to go unchecked and decreases public safety, which is a
legitimate concern and the primary motivation for \ac{EA}.

However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling identity
theft, inappropriate data mining, \ac{masssurv}, and other threats. To the extent that encryption mitigates these, it
increases public safety.

Finally, deciding how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
considerable questions of trust. Law enforcement agencies, intelligence agencies, technology makers, individual users,
and the public at large are all players; each have demonstrated untrustworthiness, and different approaches disperse
trust differently.

This complex mix of issues makes encryption policy a \ac{wicked-prob}. Even just the problem requires explanation upon
explanation and research upon research. Any formulation is concomitant with a solution, but the outcome of a solution
cannot be accurately evaluated in isolation. The underlying conditions constantly change due to external forces and
policy actions (or lack thereof).

@@

Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption does pose certain threats to public
safety. Terrorists and criminals can and do use encryption to evade the law; this has allowed for shocking levels of
\acf{CSAM} to be peddled \cite{keller_internet_2019}. However, the argument that \ac{encryption} is causing \acl{LE}
failure is not currently strong enough to justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor
coordination and training \cite{carter_2018}; for \ac{DAR}, commercial \ac{lawful-hacking} tools currently have the
upper hand on mobile phone manufacturers, meaning \acl{LE} can get into almost any mobile device they acquire
\cite{koepke_2020}. Additionally, strong \ac{encryption} is necessary to protect security and privacy---two public goods
that other technological developments often fail to provide.

Though the pro-\ac{EA} argument is ultimately not strong enough to justify an \ac{EA} mandate, incorporating it to the
technologist's point of view teaches important lessons. One is that holistic cryptosystem threat models must include two
non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and this
power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating steps.
Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption; lawmakers
represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too dangerous.
The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.
